Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-09-22 methods_paper/fabl-paper.tex
more writing.

\textbf{Our Contributions}

@@

%Most record linkage techniques are based the seminal \citep{fs}, which formalized the methods outlined by \citep{newcombe}. 
>>>>>>> e17e46b99593a9cae2676267d508dfcd21a5d0ac

@@

Several authors have elaborated on this approach by including hierarchical structure on model parameters, aiding classification through regression to incorporate covariates, establishing joint models for linkage and analysis, and more \citep{larsen2005}, \citep{hu2017}, \citep{hof2017}. Importantly, \citep{enamorado2019} modified the Fellegi-Sunter method with innovative hashing techniques, greatly expanding the scalability of the independent classification approach.

Despite the appeal of independent classification, it often leads to linkage decisions that do not enjoy transitive closure, and ignores that fact that we are often in scenarios where we know one record in one file has at most one match in another file. This setting is known as bipartite matching, and has received much less attention in the literature. \citep{sadinle_bayesian_2017} developed a prior distribution on the space of bipartite matchings such that his method, known as Beta Record Linkage (BRL), resulted in one-to-one matchings without any need for post-processing. However, BRL was even more computationally intensive than the original Fellegi-Sunter method, making it infeasible for large linkage tasks.

In this paper, we propose an extension to the \texttt{BRL} method proposed by \citep{sadinle_bayesian_2017} for this very end.  We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity. Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks. To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
pronounced ``fable'')

>>>>

\textbf{Our Contribution} 
In this paper, we propose a Bayesian extension of \cite{sadinle_bayesian_2017} (\text{BRL}) through a modification to the model specification that allows for parallel computing of the linkage parameter, using hashing techniques to hasten calculations, and to reduce computational complexity.  Specifically, \cite{sadinle_bayesian_2017} proposes a prior distribution that strictly enforces one-to-one matching. Instead, we propose independent priors (breaking the one-to-one matching) in order to create computational speeds and gains. Second, we resolve the one-to-one matching issue via a post processing step. Third, we derive conditional distributions, a Gibbs sampler, and corresponding Bayes estimates for our proposed model. Fourth, we explore other computational speeds up based upon storage efficient indexing. Finally, we explore a sensitivity analysis and revisit the case study of the El Salvadoran conflict to understand the practical ramifications regarding our proposed modeling choices. 
To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl},
pronounced ``fable'')

% We increase the speed of the \texttt{BRL} procedure through a modification to the model specification that allows for parallel computing of the linkage parameter, and use hashing techniques to hasten calculations and reduce computational complexity. Additionally, we introduce storage efficient indexing (SEI), and a data partitioning approach to make the method amenable to large linkage tasks. To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{Fast Beta Linkage} (\texttt{fabl},
%pronounced ``fable'')

@@

%Most record linkage techniques are based the seminal \citep{fs}, which formalized the methods outlined by \citep{newcombe}.

@@

%Several authors have elaborated on this approach by including hierarchical structure on model parameters, aiding classification through regression to incorporate covariates, establishing joint models for linkage and analysis, and more \citep{larsen2005}, \citep{hu2017}, \citep{hof2017}. Importantly, \citep{enamorado2019} modified the Fellegi-Sunter method with innovative hashing techniques, greatly expanding the scalability of the independent classification approach. 
%
%Despite the appeal of independent classification, it often leads to linkage decisions that do not enjoy transitive closure, and ignores that fact that we are often in scenarios where we know one record in one file has at most one match in another file. This setting is known as bipartite matching, and has received much less attention in the literature. \citep{sadinle_bayesian_2017} developed a prior distribution on the space of bipartite matchings such that his method, known as Beta Record Linkage (BRL), resulted in one-to-one matchings without any need for post-processing. However, BRL was even more computationally intensive than the original Fellegi-Sunter method, making it infeasible for large linkage tasks.
